---
title: "behaviourFam3"
author: "Karissa Barthelson"
date: "2024-07-23"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  autodep = TRUE,
  echo = TRUE, 
  warning = FALSE,
  message = FALSE,
  fig.align = "center", 
  fig.retina = 1,
  out.width ="100%", 
  out.height = "100%"
)
```

```{r packages}
library(tidyverse)
library(readxl)
library(magrittr)

# vis
library(scales)
library(ggpubr)
library(ggeasy)
library(ggfortify)
library(ggbeeswarm)
library(ggforce)
library(ggrepel)
library(kableExtra)
library(pander)

# stat analysis
library(broom)
library(lme4)
library(performance)
library(car)
library(emmeans)
library(glmmTMB)
library(MASS)

# set the default theme for ggplot as theme_bw
theme_set(
  theme_bw() + 
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold")
    ))


```

# Introduction and data import

In the previous analyses, I did not have balanced sexes in the behaviour testing as I did not realise sex was an important descriptor. Therefore, I repeated the iron treatments and behaviour experiment on a new fam of fish which had balanced sexes (20 females and 20 males per sex and treatment)


```{r importMeta}
# read in metadata and cleanup cols
meta <- 
  read_excel("data/ymaze/fam3/20240711 naglu hom x het 6mo FeCit_fam3_metadata.xlsx") %>% 
  mutate(
    
    # convert to factors
    fish_id = as.factor(fish_id), 
    HomeTank = as.factor(HomeTank), 
    treatment = fct_inorder(treatment), 
    sex = as.factor(sex), 
    behavBatchID = as.factor(behavBatchID), 
    genotype = factor(genotype, c("het", "hom")), 
  
  # create some new variables which are helpful for the 
  geno_treat = paste0(genotype, "_", treatment) %>% as.factor(),
  sex_geno_treat = paste0(sex, "_", genotype, "_", treatment) %>% as.factor()
  ) 
```

### number of fish

This time, the number of fish across groups is much more balanced. Although there were less male homs treated with iron - this is just the luck of the draw in being blinded during the experiment. 

```{r numfishplot}
meta %>% 
  ggplot(
    aes(x = geno_treat, fill = sex)
  ) +
  geom_bar(position = "dodge") +
  scale_y_continuous(
    breaks = 0:15
  )+
  facet_wrap(~sex) +
  labs(
    title = "Number of fish in the experiment", 
    x = "Experimental Group"
  ) +
  theme(axis.text.x = element_text(angle = -45, hjust = 0))
```

# explanation of processing videos 

## concat & compress
For each behav trial, there were 2 videos produced. This is because Zantiks has a max file size of 2Gb. These raw videos in the `data/ymaze/fam3/raw_videos/` folder were first concatenated using ffmpeg to make a single video. They were then compressed so the file size was smaller. This processing was performed using the `concat videos.py` in the scripts folder. 

```{r concatvidsScript, attr.output='style="max-height: 150px;"'}
readLines("data/ymaze/fam3/scripts/python/concat videos.py")
```

## splitting, bg substration. 
The next step was splitting the videos into the individual Y-shapes, as this is easier for tracking by sleap.ai. 

First, needed to extract a frame using ffmeg. This was performed using this script. 

```{r bashextractframe, attr.output='style="max-height: 150px;"'}
readLines("data/ymaze/fam3/scripts/sh/extract_frames.bash")
```

These frames were edited in GIMP to make masks for video splitting.

The videos were bg substracted, gaussian blurred and split to individual Y mazes using this script below. This took hours on my laptop. 

```{r videopreprocessing, attr.output='style="max-height: 150px;"'}
readLines("data/ymaze/fam3/scripts/python/KB local video preprocessing.py")
```

The videos are now ready for tracking using sleap.ai. 

# details of sleap.ai tracking

Checked it installed using this command

```conda activate sleap```

```sleap-label```


Labelled 31 frames with the head, body and tail of a female fish (fish)




