---
title: "metabolomics"
author: "Karissa Barthelson"
date: "2024-03-13"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r loadLibs}
# data wrangling
library(tidyverse)
library(magrittr)
library(readxl)
library(qs)

# bioconductor
library(msigdbr)
library(MetaboAnalystR)
library(KEGGREST)
library(limma)
library(msigdbr)

# vis
library(pheatmap)
library(ggpubr)
library(ggrepel)
library(ggeasy)
library(ggfortify)
library(ggforce) # for facet_zoom
library(RColorBrewer)
library(UpSetR)
library(pander)
library(ggeasy)
library(scales)
library(patchwork)

colours =  c("grey20", "grey80", "salmon", "seagreen2" )

# set default ggplot theme
theme_set(theme_classic())
```


```{r install}
# function to download the dependencies
# metanr_packages <- function(){
# metr_pkgs <- c("impute", "pcaMethods", "globaltest", "GlobalAncova", "Rgraphviz", "preprocessCore", "genefilter", # "SSPA", 
#                "sva", "limma", "KEGGgraph", "siggenes","BiocParallel", "MSnbase", "multtest", "RBGL", "edgeR", "fgsea", "devtools", "crmn")
# list_installed <- installed.packages()
# new_pkgs <- subset(metr_pkgs, !(metr_pkgs %in% list_installed[, "Package"]))
# if(length(new_pkgs)!=0){if (!requireNamespace("BiocManager", quietly = TRUE))
#         install.packages("BiocManager")
#         BiocManager::install(new_pkgs)
#         print(c(new_pkgs, " packages added..."))
#     }
# 
# if((length(new_pkgs)<1)){
#         print("No new packages added...")
#     }
# }
# 
# metanr_packages()
# 
# library(devtools)
# # Step 2: Install MetaboAnalystR with documentation
# devtools::install_github("xia-lab/MetaboAnalystR", 
#                          build = TRUE, 
#                          build_vignettes = TRUE, 
#                          build_manual =TRUE)
```


# import data as a Mset object

There are no missing points so will not run the missing objects code. 

```{r}
setwd("~/analyses/MPS III iron supplementation")
# make a metadata object
meta <- 
  read_excel("data/metabolomics/FINAL_Karissa Barthelson_LCMS_MA-KB-268/Delivered Data_xlsx&ppt/MA-KB-268_LCMS_Matrix copy.xlsx") %>% 
  dplyr::select(1) %>% 
  mutate(group = str_extract(
    Sample, pattern = "h.+_"
  ), 
  group = str_remove_all(group, pattern = "_"))

setwd("data/metabolomics/")
# iniitalise the mSet object
mSet <- InitDataObjects("pktable", "stat", FALSE)

# add the peak intensities 
mSet <- Read.TextData(
  mSet, 
  filePath = "FINAL_Karissa Barthelson_LCMS_MA-KB-268/Delivered Data_xlsx&ppt/MA-KB-268_LCMS_Matrix copy.csv", 
  format = "rowu", 
  lbl.type = "disc"
)
# perform a "Sanity check" which looks at sample and class labels, looks at NA vals etc. see ?SanityCheckData
mSet<-SanityCheckData(mSet)

# filter based off RSD (RSD = SD/mean)
# maybe wont do this/ 
#mSet <- FilterVariable(mSet, "F", 25, "rsd", 20)
```


```{r}

# normalisation, log transformation adn mean/center scaling
mSet <- ReplaceMin(mSet)
mSet <- PreparePrenormData(mSet)

mSet <- 
  Normalization(
    mSet, 
    rowNorm = "MedianNorm", # median normalisation
    transNorm =  "LogNorm",  # log transformation  
    scaleNorm = FALSE # dont center and scale

  )   
```

## distributions

Samples are all on the same scale and are comparable. 

### sample level
```{r}
mSet$dataSet$norm %>% 
  rownames_to_column("Sample") %>% 
  as_tibble() %>% 
  dplyr::filter(
    # !grepl(Sample, pattern = "QC")
  ) %>%
  left_join(meta) %>% 
  pivot_longer(names_to = "metabolite", values_to = "intensity.normalised", contains("_")) %>% 
  ggplot(
    aes(y = intensity.normalised, x = Sample, fill = group)
    )+
  geom_boxplot() +
  theme(
    axis.text.x = element_text(angle = -45, hjust = 0), 
    plot.margin = margin(t = 1, l = 1, b = 1, r = 4, unit = "cm")
  )

```

### metabolite level

```{r}
mSet$dataSet$norm %>% 
  rownames_to_column("Sample") %>% 
  as_tibble() %>% 
  dplyr::filter(
    # !grepl(Sample, pattern = "QC")
  ) %>%
  left_join(meta) %>% 
  pivot_longer(
    names_to = "metabolite", 
    values_to = "intensity.normalised", contains("_")
    ) %>% 
  ggplot(
    aes(y = intensity.normalised, x = metabolite)
    )+
  geom_boxplot() +
  theme(
    axis.text.x = element_text(angle = -45, hjust = 0), 
    plot.margin = margin(t = 1, l = 1, b = 1, r = 4, unit = "cm")
  )
```


# PCA 

PCA was performed on all samples (not the QCs). 2 pretty clear outliers are present. maybe will remove them.
```{r}
mSet$dataSet$norm %>% 
  rownames_to_column("sample") %>% 
  as_tibble() %>% 
  dplyr::filter(
    !grepl(sample, pattern = "QC"),
    !grepl(sample, pattern = "EB")
  ) %>%
  column_to_rownames("sample") %>% 
  prcomp() %>%
  autoplot(
    data = tibble(Sample = rownames(.$x)) %>%
      left_join(meta),
    colour = "group", 
    size = 6
  ) +
  scale_color_manual(
    values = c("grey20", "grey80", "salmon", "seagreen2" )
  ) +
  theme(
    legend.position = "bottom", 
    aspect.ratio = 1
  ) +
  ggsave(
    "output/plots/PCA_metabolome.png", 
    width = 11, height = 10, units = "cm", 
    scale = 1.5, 
    dpi = 400
  )
```

PCA was repeated omitting the het iron group. Really nice sample grouping is observed. Samlples seperate by treatment across PC1, and genotype by PC1 and PC2. 

```{r}
mSet$dataSet$norm %>% 
  rownames_to_column("sample") %>% 
  as_tibble() %>% 
  dplyr::filter(
    !grepl(sample, pattern = "QC"),
    !grepl(sample, pattern = "EB"), 
    !grepl(sample, pattern = "het_10")
  ) %>% 
  column_to_rownames("sample") %>% 
  prcomp() %>%
  autoplot(
    data = tibble(Sample = rownames(.$x)) %>%
      left_join(meta),
    colour = "group", 
    size = 6
  ) 
  # stat_ellipse(
  #   aes(
  #     group = group, 
  #     colour = group, 
  #     fill = group), 
  #   ) +
```




```{r}
normalised.data <- mSet$dataSet$norm %>% 
  rownames_to_column("sample") %>% 
  as_tibble() %>% 
  dplyr::filter(
    !grepl(sample, pattern = "QC"),
    !grepl(sample, pattern = "EB")
  ) %>%
  column_to_rownames("sample") %>% 
  t() %>% 
  as.data.frame() %>% 
  rownames_to_column("temp") %>% 
  as_tibble() %>% 
  separate(temp, # seperate the name and id into seperate cols
           into = c("name", "id"), 
           sep = "_") %>% 
  pivot_longer(
    names_to  = "sample", 
    values_to = "intensity.normalised", 
    starts_with("KB"))
```

<!-- # PLS-DA -->
<!-- ```{r} -->
<!-- mSet <- PLSR.Anal(mSet, reg=TRUE)  -->
<!-- mSet <-  PlotPLSPairSummary(mSet, pc.num = 5,  -->
<!--                      imgName = "pls_pair_0_", format = "png",  -->
<!--                      dpi = 72, width=NA) -->


<!-- mSet <- PlotPLS2DScore(mSet, imagen"pls_score2d_0_", "png", 72, width=NA, 1,2,0.95,0,0, "na") -->

<!-- 37.	mSet<-PlotPLS2DScore(mSet, "pls_score2d_0_", "png", 72, width=NA, 1,2,0.95,0,0, "na") -->
<!-- 38.	mSet<-PlotPLS3DScoreImg(mSet, "pls_score3d_0_", "png", 72, width=NA, 1,2,3, 40) -->
<!-- 39.	mSet<-PlotPLSLoading(mSet, "pls_loading_0_", "png", 72, width=NA, 1, 2); -->
<!-- 40.	mSet<-PlotPLS3DLoading(mSet, "pls_loading3d_0_", "json", 1,2,3) -->
<!-- 41.	mSet<-PlotPLS.Imp(mSet, "pls_imp_0_", "png", 72, width=NA, "vip", "Comp. 1", 15,FALSE) -->
<!-- ``` -->

# limma for differential expresison 

```{r}
# prepare a design matrix
design <- model.matrix(~0 + group, 
                       data = meta) %>% 
  set_colnames(
    colnames(.) %>% 
      str_remove(pattern = "group")
    )

# set contrasts of interest
contrasts <- 
  makeContrasts(
    hom.v.het = hom0.85saline - het0.85saline, 
    hetiron.v.hetsaline = het10ugFeCit - het0.85saline, 
    homiron.v.homsaline = hom10ugFeCit - hom0.85saline,
    homiron.v.hetsaline = hom10ugFeCit - het0.85saline,
    
    levels = colnames(design)
    )

# prepare the data forinput
limma.input <- normalised.data %>% 
  pivot_wider(names_from = "sample", 
              values_from = "intensity.normalised") %>% 
  column_to_rownames("name") %>% 
  dplyr::select(-id)

# fit linear model 
fit <- lmFit(
  object = limma.input, 
  design = design
)

fit.contrasts <- colnames(contrasts) %>% 
  sapply(function(x) {
    fit %>% 
      contrasts.fit(
        contrasts = contrasts[,x]
      ) %>% 
      eBayes(
        # trend=TRUE, 
    robust=TRUE
  )
  }, simplify = F)

# call the toptables

toptable_1 <- 
  fit.contrasts %>% 
  lapply(function(x) {
    x %>% 
      topTable(n = Inf) %>% 
      as.data.frame() %>% 
      rownames_to_column("name") %>% 
      left_join(normalised.data %>% 
                  dplyr::select(name, id) %>% unique) %>% 
      as_tibble %>% 
      dplyr::rename(
        FDR = adj.P.Val
      )
  })

# add names of the coef. 
mynames <- names(toptable_1)
toptable_1 <- map2(toptable_1, mynames,
                   ~.x %>% mutate(coef = .y))

```

# data vis
```{r}
# volcano plot
toptable_1 %>% 
  bind_rows() %>% 
  ggplot(
    aes(x = logFC, y = -log10(P.Value), 
        colour = FDR < 0.05), 
    ) +
  geom_point() +
  scale_color_manual(
    values = c("grey50", "red")
  ) +
  geom_label_repel(
    aes(label = name), 
    show.legend = FALSE, 
    data = . %>% 
      dplyr::filter((logFC > 0.4 | logFC < -0.4))
    ) +
  facet_wrap(~coef)
```

# boxplots of top 10 DE metabolites
```{r}
normalised.data %>% 
  dplyr::filter(name %in% (toptable_1$hom.v.het %>% head(6) %>% .$name)) %>% 
  left_join(meta %>% dplyr::rename(sample = Sample)) %>% 
   dplyr::filter(
     !grepl(group, pattern = "het10ugFeCit")
     ) %>%
  ggplot(
    aes(x = group, y = intensity.normalised)
    ) +
  geom_boxplot(
    aes(fill = group),
    outlier.shape = NA
    ) +
  geom_jitter() +
  facet_wrap(
    ~name, 
    scale = "free"
    ) +
  scale_fill_manual(
    values = colours[-2]
    ) +
  theme(
    axis.text.x = element_text(angle = -45, hjust = 0, vjust = 1), 
    legend.position = "none", 
    plot.margin = margin(t = 1, l = 1, r = 2, unit = "cm")
    )

```

# Pathway analysis 

I next want to see how the metabolites are changing across KEGG pathways. 

`MetaboAnalystR` creates a file called "compound_db.qs". This contains the mapping between HMDB and KEGG ids. I will use this to make the object for gene set testing. 

```{r}
compounds <- qread("compound_db.qs") %>% 
  as_tibble %>% 
  dplyr::filter(hmdb_id %in% normalised.data$id) %>% 
  dplyr::rename(kegg_compound_id = kegg_id)

KEGG.pw2cp <- keggLink("pathway", "compound") %>% 
  unlist() %>% 
  tibble(kegg_pathway_id = ., 
         kegg_compound_id = sub("cpd:", "", names(.))) %>% 
  mutate(
    kegg_pathway_id = str_remove(kegg_pathway_id, 
                                 pattern = "path:"))

KEGG.pwid2pw <- keggList("pathway") %>% 
  unlist %>% 
  tibble(kegg_pathway_name = ., 
         kegg_pathway_id = sub("cpd:", "", names(.)))

KEGG <- compounds %>% 
  left_join(KEGG.pw2cp) %>% 
  left_join(KEGG.pwid2pw) %>% 
  split(f = .$kegg_pathway_name) %>% 
  lapply(function(x) {
    x$hmdb_id
  })

# omit gene sets with only 2 metabolitea
KEGG <- KEGG[lapply(KEGG, length) > 2]
```

## rotation gene set testing 

Using `limma:fry` no gene sets reached statistical significance
```{r}
fry(normalised.data %>% 
      pivot_wider(names_from = "sample", 
                  values_from = "intensity.normalised") %>% 
      column_to_rownames("id") %>% 
      dplyr::select(-name), 
    index = KEGG, 
    design = design, 
    sort = "mixed",
    contrast = contrasts) %>% 
  rownames_to_column("kegg") %>% 
  as_tibble
```

## GSEA
I also want to try GSEA. Metabolites were ranked on the t- statistic. One gene set was found to be statistically significant 

```{r}
ranks <- setNames(toptable_1$hom.v.het$t, toptable_1$hom.v.het$id)

# set a seed for a reproducible result 
set.seed(1)
fgseaRes <- 
  fgsea::fgseaMultilevel(
  pathways = KEGG, stats = ranks, ) %>% 
  as_tibble() %>% 
  mutate(padj = p.adjust(pval, method = "fdr")) %>% 
  arrange(padj)

fgseaRes
```

## plot the results
```{r}
anno <- meta %>% 
  mutate(
    genotype = case_when(
      grepl(group, pattern = "het") ~ "het", 
      grepl(group, pattern = "hom") ~ "hom"
    ), 
    treatment = case_when(
      grepl(group, pattern = "salin") ~ "0saline", 
      grepl(group, pattern = "Fe") ~ "1Fe"
    )
  ) %>% 
  dplyr::select(-group) %>% 
  column_to_rownames("Sample")

ann_colors <- list(
    genotype = c(het = "white",hom = "firebrick"),
    treatment = c(`0saline` = "grey50", `1Fe` = "#D95F02")
)

```


### metabolite abundances
### heatmaps
```{r}
normalised.data %>% 
  dplyr::filter(id %in% fgseaRes[1,]$leadingEdge[[1]]) %>% 
  dplyr::select(name, intensity.normalised, sample) %>% 
  pivot_wider(names_from = sample, 
              values_from = intensity.normalised) %>% 
  column_to_rownames("name") %>% 
  pheatmap(
    scale = "row", 
    annotation_col = anno, annotation_colors = ann_colors,
    color = colorRampPalette(rev(brewer.pal(n = 7, name =
  "RdBu")))(100), 
  main = "Valine, leucine and isoleucine biosynthesis"
  )
```

```{r}
normalised.data %>% 
  dplyr::filter(id %in% fgseaRes[2,]$leadingEdge[[1]]) %>% 
  dplyr::select(name, intensity.normalised, sample) %>% 
  pivot_wider(names_from = sample, 
              values_from = intensity.normalised) %>% 
  column_to_rownames("name") %>% 
  pheatmap(
    scale = "row", 
    annotation_col = anno, annotation_colors = ann_colors,
    color = colorRampPalette(rev(brewer.pal(n = 7, name =
  "RdBu")))(100), 
  main = "Valine, leucine and isoleucine degradation"
  )
```

## boxplots
```{r}
normalised.data %>% 
  dplyr::filter(id %in% fgseaRes[1,]$leadingEdge[[1]]) %>% 
  dplyr::select(name, intensity.normalised, sample) %>%
  left_join(anno %>% rownames_to_column("sample")) %>% 
  ggplot(
    aes(x = treatment, y = intensity.normalised, 
        colour = genotype)
  ) +
  geom_boxplot(
      outlier.shape = NA
      ) + 
  geom_point(
    position = position_jitterdodge()
  ) +
  facet_wrap(~name)


normalised.data %>% 
  dplyr::filter(id %in% fgseaRes[2,]$leadingEdge[[1]]) %>% 
  dplyr::select(name, intensity.normalised, sample) %>%
  left_join(anno %>% rownames_to_column("sample")) %>% 
  ggplot(
    aes(x = treatment, y = intensity.normalised, 
        colour = genotype)
  ) +
  geom_boxplot(
      outlier.shape = NA
      ) + 
  geom_point(
    position = position_jitterdodge()
  ) +
  facet_wrap(~name)
```


# Power calculation

I'm not sure the R package can do the power calculations. So what I did was perform it in the metaboanalast website. I then downloaded the output.

```{r}

p1 <- normalised.data %>%
  dplyr::filter(name %in% c("myo-Inositol",  "Erucic acid"   )) %>%
  left_join(meta %>% dplyr::rename(sample = Sample)) %>%
   dplyr::filter(
     !grepl(group, pattern = "FeCit")
     ) %>%
  mutate(
    group = case_when(
      grepl(group, pattern = "het") ~ "het",
      grepl(group, pattern = "hom") ~ "MPS IIIB "
  )) %>%
  ggplot(
    aes(x = group, y = intensity.normalised)
    ) +
  geom_boxplot(
    aes(fill = group),
    outlier.shape = NA
    ) +
  geom_jitter() +
  ggpubr::stat_compare_means(
    label = "p",
    label.x = 1.5,
    method = "t.test"
    ) +
  facet_wrap(
    ~name,
    scale = "free"
    ) +
  scale_fill_manual(
    values = c(viridis_pal()(4)[2],
               "grey50",
               viridis_pal()(4)[3:4]
    )
  ) +
  labs(
    title = "A  MPS IIIB"
  ) +
  theme(
    text = element_text(face = "bold", size = 18),
    legend.position = "none"
  )

p3 <- read_csv("Power calc download/predicted_sample_powers.csv") %>%
  ggplot(aes(
    x = `Sample Size (per group)`,
    y = `Predicted power`
  )) +
  geom_point() +
  geom_line() +
  geom_hline(yintercept = 0.7) +
  scale_x_continuous(limits = c(0,25),
                     breaks = seq(0,25, by = 5)) +
  scale_y_continuous(breaks = seq(0,1, by = 0.1)) +
  labs(
    title = "C  Power calculation"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    text = element_text(face = "bold", size = 18)
  )

p2 <- p2 +
  ggtitle("B   EOfAD")

  (p1 + p2)/p3 +
  ggsave("metabolitesforIdeas.png",
         width = 12, height = 10, units = "cm", dpi = 300, scale = 2)
```
# export
```{r}
toptable_1$hom.v.het %>% 
  write_csv("naglu toptable.csv")

normalised.data %>% 
  saveRDS("naglu normalised data.rds")
```

