---
title: "WGCNA_comp_consens_with_individ_nets"
author: "Karissa Barthelson"
date: "2024-06-20"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

# setup

Load packages

```{r}
library(tidyverse)
library(magrittr)
library(readxl)

# vis
library(pander)
library(kableExtra)
library(UpSetR)
library(scales)
library(pheatmap)
library(ggpubr)
library(ggrepel)
library(ggfortify)
library(ggforce)
library(RColorBrewer)
library(colorspace)
library(igraph)
library(ggvenn)

# analysis
library(WGCNA)
library(topGO)
library(org.Dr.eg.db)
library(clusterProfiler)
library(KEGGREST)
library(msigdbr)
library(matrixStats)

theme_set(theme_classic())
options(stringsAsFactors = FALSE)
allowWGCNAThreads()  
```

# Import consensus and female data

Here, I import the consensus modules (multi-expr object) and the female specific modules.

```{r}
# load the consensus data
load("output/WGCNA/R_objects/Consensus-NetworkConstruction-man.RData")

# rename moduleColours to be the consensus ones, remove modulecolours to save any confusion
consColours = moduleColors
remove(moduleColors)

# also do this for the moduleLabels
cons.module.Labels = moduleLabels
remove(moduleLabels)

# load the female specific network objects
mergedColors.signed.females <- 
  readRDS("output/WGCNA/R_objects/mergedColors.signed.females.rds")
merge.signed.females <- 
  readRDS("output/WGCNA/R_objects/merge.signed.females.rds")
```

According to a bioconductor support [post](https://support.bioconductor.org/p/129821/) the female and consensus modules need to be restricted to the same genes. So import the gene ids in each, and find the common ones. First, I check whether the same genes are being used in each network, which is mostly true. 

```{r}
# get female genes from the adjacency matrix rownames
femGenes <-
  readRDS("output/WGCNA/R_objects/adjacency.signed.females.rds") %>% 
  rownames()

# consensus genes from the multi-expr object
consGenes = mtd.colnames(multiExpr)

# check the overlap
ggvenn::ggvenn(
  list(
    female.genes = femGenes, 
    consensus.genes = consGenes
  )
)
```


```{r}
# find the intersection
common = intersect(femGenes, consGenes)

# only keep the common genes in the modules/ 
femaleColors = mergedColors.signed.females[match(common, femGenes)]
consColours.common = consColours[match(common, consGenes)]
```


```{r}
# Rename variables to avoid conflicts
femaleMEs = orderMEs(merge.signed.females$newMEs, greyName = "ME0")
```

The consensus network analysis results are represented by the variables `consMEs`, `moduleLabels`, `moduleColors`, and `consTree.` We are now ready to relate the female modules to the consensus modules. We calculate the overlaps of each pair of female-consensus modules, and use the Fisher’s exact test (also known as hypergeometric test) to assign a p-value to each of the pairwise overlaps.

```{r}
# Isolate the module labels in the order they appear in ordered module eigengenes
# remove the ME from each name
femModuleLabels = substring(names(femaleMEs), first =  3)
consModuleLabels = substring(names(consMEs[[1]]$data),first =  3)

# Convert the numeric module labels to color labels
femModules = femModuleLabels
consModules = labels2colors(consModuleLabels)

# Numbers of female and consensus modules
nFemMods = length(femModules)
nConsMods = length(consModules)

# Initialize tables of p-values and of the corresponding counts
pTable = matrix(0, nrow = nFemMods, ncol = nConsMods)
CountTbl = matrix(0, nrow = nFemMods, ncol = nConsMods)

# Execute all pairwise comparisons
for (fmod in 1:nFemMods)
for (cmod in 1:nConsMods)
{
femMembers = (femaleColors == femModules[fmod])
consMembers = (consColours.common == consModules[cmod])

pTable[fmod, cmod] = -log10(fisher.test(
  femMembers, consMembers, alternative = "greater")$p.value)

CountTbl[fmod, cmod] = sum(femaleColors == femModules[fmod] & 
                             consColours.common == consModules[cmod])
}
```

To display the p-value and count tables in an informative way, we create a color-coded table of the intersection counts. The colors will indicate the p-value significance:

```{r}
# Truncate p values smaller than 10^{-50} to 10^{-50}
pTable[is.infinite(pTable)] = 1.3*max(pTable[is.finite(pTable)]);
pTable[pTable>50 ] = 50 ;

# Marginal counts (really module sizes)
femModTotals = apply(CountTbl, 1, sum)
consModTotals = apply(CountTbl, 2, sum)

# Actual plotting
sizeGrWindow(15,7)
par(mfrow=c(1,1))
par(cex = 1.0)
par(mar=c(10,14,4,0)+2) 

labeledHeatmap(
  Matrix = pTable,
  xLabels = paste(" ", consModules),
  yLabels = paste(" ", femModules),
  
  colorLabels = TRUE,
  
  xSymbols = paste("Cons ", consModules, ": ", consModTotals, sep=""),
  ySymbols = paste("Fem ", femModules, ": ", femModTotals, sep=""),
  
  textMatrix = CountTbl,
  
  colors = blueWhiteRed(100)[50:100],
  
  main = "Correspondence of Female set-specific and Female-Male consensus modules",
  
  cex.text = 1.0, cex.lab = 1.0, 
  setStdMargins = FALSE,
  plotLegend = TRUE
  )
```

Each row of the table corresponds to one female set-specific module (labeled by color as well as text), and each column corresponds to one consensus module. Numbers in the table indicate gene counts in the intersection of the corresponding modules. Coloring of the table encodes − log(p), with p being the Fisher’s exact test p-value for the overlap of the two modules. The stronger the red color, the more significant the overlap is. The table indicates only one female set-specific modules have a consensus counterpart. This is the female royalblue module.

To remind myself, I then looked into the over-representation of GO terms in genes in this module. See the table below.

```{r}
enrichment_table_female_GOterms <- 
  readRDS("/output/WGCNA/enrichment/enrichment_table_female_GOterms.rds")

enrichment_table_female_GOterms %>% 
  dplyr::filter(module == "royalblue") %>% 
  kable %>% 
  kable_styling()
```

# Male data

I now want to look at whether the male modules have a consensus module.

```{r}
# load the female specific network 
mergedColors.signed.males <- 
  readRDS("output/WGCNA/R_objects/mergedColors.signed.males.rds")
merge.signed.males <- 
  readRDS("output/WGCNA/R_objects/merge.signed.males.rds")

# Rename to moduleColors
maleColors = mergedColors.signed.males

# Construct numerical labels corresponding to the colors
colorOrder = c("grey", standardColors(28))
maleLabels = match(moduleColors, colorOrder)-1

maleMEs = merge.signed.males$newMEs
maleMEs = orderMEs(maleMEs, greyName = "ME0")


# Isolate the module labels in the order they appear in ordered module eigengenes
maleModuleLabels = substring(names(maleMEs), 3)
#consModuleLabels = substring(names(consMEs[[1]]$data), 3)

# Convert the numeric module labels to color labels
maleModules = labels2colors(maleModuleLabels)
#consModules = labels2colors(consModuleLabels)

# Numbers of female and consensus modules
nMaleMods = length(maleModules)
nConsMods = length(consModules)

# Initialize tables of p-values and of the corresponding counts
pTable.males = matrix(0, nrow = nMaleMods, ncol = nConsMods)
CountTbl.males = matrix(0, nrow = nMaleMods, ncol = nConsMods)

# Execute all pairwaise comparisons
for (fmod in 1:nMaleMods)
for (cmod in 1:nConsMods)
{
maleMembers = (maleColors == maleModules[fmod])

consMembers = (moduleColors == consModules[cmod])

pTable.males[fmod, cmod] = -log10(fisher.test(
  maleMembers, consMembers, alternative = "greater")$p.value)

CountTbl.males[fmod, cmod] = sum(maleColors == maleModules[fmod] & moduleColors ==
consModules[cmod])
}

# Truncate p values smaller than 10^{-50} to 10^{-50}
pTable.males[is.infinite(pTable.males)] = 1.3*max(pTable.males[is.finite(pTable.males)])

pTable.males[pTable.males>50 ] = 50 

# Marginal counts (really module sizes)
maleModTotals = apply(CountTbl.males, 1, sum)
consModTotals = apply(CountTbl.males, 2, sum)

# Actual plotting
sizeGrWindow(15,7)
par(mfrow=c(1,1))
par(cex = 1.0)
par(mar=c(10,14,4,0)+2) 

labeledHeatmap(
  Matrix = pTable.males,
  xLabels = paste(" ", consModules),
  yLabels = paste(" ", maleModules),
  
  colorLabels = TRUE,
  
  xSymbols = paste("Cons ", consModules, ": ", consModTotals, sep=""),
  ySymbols = paste("male ", maleModules, ": ", maleModTotals, sep=""),
  
  textMatrix = CountTbl.males,
  
  colors = blueWhiteRed(100)[50:100],
  
  main = "Correspondence of Male set-specific and Female-Male consensus modules",
  
  cex.text = 1.0, cex.lab = 1.0, 
  setStdMargins = FALSE,
  plotLegend = TRUE
  )
```

Only one male module has a representative in the consensus modules.

```{r}
enrichment_table_male_GOterms <- 
  readRDS("output/WGCNA/enrichment/enrichment_table_male_GOterms.rds")

enrichment_table_male_GOterms %>% 
  dplyr::filter(module== "lightyellow") %>% 
  kable %>% 
  kable_styling()
```
